---
title: "SoilGrids"
format: html
---

```{r}
#| label: setup
#| include: false

source(here::here("R/quarto-setup.R"))
```

## Overview

This document provides a mean to download, import and plot the data from the Global Gridded Soil Information ([SoilGrids ](https://www.isric.org/explore/soilgrids)) database.

## Setting things up

```{r}
#| eval: false

library(beepr)
library(dplyr)
library(ggplot2)
library(magrittr)
library(terra)
library(tidyterra)
library(spData)
library(viridis)
```

```{r}
#| include: false

library(magrittr)
library(terra)
library(tidyterra)
```

## Downloading data

For more information, see: 

- https://www.isric.org/explore/soilgrids
- https://soilgrids.org
- https://www.isric.org/explore/soilgrids/soilgrids-access
- https://git.wur.nl/isric/soilgrids/soilgrids.notebooks/-/blob/master/markdown/webdav_from_R_terra.md
- https://git.wur.nl/isric/soilgrids/soilgrids.notebooks/-/blob/master/markdown/webdav_from_R.md

- Variable: Nitrogen.
- Depth: 0-5.
- Value: Mean in cg/kg.
- Spatial resolution: 5000m cell size.

```{r}
source <- "https://files.isric.org/soilgrids/latest/data/nitrogen/nitrogen_0-5cm_mean.vrt"
```

```{r}
resolution <- "5000m" # or "1000m"" or "5000m"

voi <- "nitrogen" # variable of interest
depth <- "0-5cm" # 0-5cm, 5-15cm, 15-30cm, 30-60cm, 60-100cm or 100-200cm
quantile <- "mean" # mean, Q0.05, Q0.5, Q0.95 or uncertainty

voi_layer <- paste(voi, depth, quantile, sep="_") # layer of interest

ras_crs <- "ESRI:54052" # ESRI code of the interrupted Goode Homolosine projection used by SoilGrids
```

```{r}
#| eval: false
#| output: false

sa_shape <- 
  # spData::world |>
  sf::st_read(system.file("shapes/world.gpkg", package="spData")) |>
  dplyr::filter(continent == "South America")
```

```{r}
# Old code.

# sa_shape |> sf::st_crs()
# sa_shape <- sa_shape |> sf::st_transform("EPSG:4326")
```

```{r}
shape_bbox <-
  sa_shape |>
  magrittr::extract2("geom") |>
  sf::st_bbox()

shape_bbox
```

```{r}
in_crs <- terra::crs(sa_shape)
```

```{r}
if (resolution == "250m") {
  rstFile <- paste0(
    "/vsicurl/https://files.isric.org/soilgrids/latest/data/",
    voi,
    "/",
    voi_layer,
    ".vrt"
  )
} else if (resolution == "1000m") {
  rstFile <- paste0(
    "/vsicurl/https://files.isric.org/soilgrids/latest/data_aggregated/1000m/",
    voi,
    "/",
    voi_layer,
    "_1000.tif"
  )
} else {
  rstFile <- paste0(
    "/vsicurl/https://files.isric.org/soilgrids/latest/data_aggregated/5000m/",
    voi,
    "/",
    voi_layer,
    "_5000.tif"
  )
}
```

```{r}
ras <- terra::rast(rstFile)

# Assign crs as the ESRI code rather than proj string.
terra::crs(ras) <- ras_crs
```

```{r}
# ras <- raster::raster(rstFile)
```

```{r}
# Project bounding box to same CRS as raster layer.
bb_proj <- terra::project(
  x = terra::ext(sa_shape),
  from = terra::crs(in_crs),
  to = terra::crs(ras_crs)
)
```

```{r}
terra::window(ras) <- NULL # Remove any existing window.

terra::window(ras) <- bb_proj # Get just ROI.
```

## Plotting the data

```{r}
#| eval: false
#| include: false

# terra::plot(ras) # Quick visual check.
```

```{r}
#| warning: false

ggplot2::ggplot() + 
  tidyterra::geom_spatraster(data = ras) +
  # ggplot2::scale_fill_distiller(
  #   palette = "Greens", 
  #   direction = 1,
  #   na.value = NA
  # ) +
  viridis::scale_fill_viridis(
    option = "D", 
    direction = 1,
    na.value = NA
  ) +
  ggplot2::theme(
    axis.title.x = ggplot2::element_blank(),
    axis.title.y = ggplot2::element_blank()
  ) +
  ggplot2::labs(fill = "")
```

## References {.smaller}

::: {style="font-size: 0.75em;"}
In accordance with the [American Psychological Association (APA) Style](https://apastyle.apa.org/), 7th edition.
:::

::: {#refs style="font-size: 0.75em;"}
:::
