---
title: "WorldClim"
format: html
---

```{r}
#| label: setup
#| include: false

source(here::here("R/quarto-setup.R"))
```

## Overview

This document provides a mean to download, import and plot the data from the [WorldClim 2.1](https://worldclim.org/) database.

## Setting things up

```{r}
#| eval: false

library(beepr)
library(curl)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(magrittr)
library(raster)
library(spData)
library(viridis)
```

```{r}
#| include: false

library(magrittr)
library(rlang)
```

## Downloading data

- GCM: ACCESS-CM2.
- Resolution: 5 arc-minutes.
- Variable: tx - monthly average maximum temperature (°C).

```{r}
sources <- list(
  ssp126 = list(
    source = "https://geodata.ucdavis.edu/cmip6/5m/ACCESS-CM2/ssp126/wc2.1_5m_tmax_ACCESS-CM2_ssp126_2021-2040.tif",
    file = tempfile()
  ),
  ssp245 = list(
    source = "https://geodata.ucdavis.edu/cmip6/5m/ACCESS-CM2/ssp245/wc2.1_5m_tmax_ACCESS-CM2_ssp245_2021-2040.tif",
    file = tempfile()
  ),
  ssp370 = list(
    source = "https://geodata.ucdavis.edu/cmip6/5m/ACCESS-CM2/ssp370/wc2.1_5m_tmax_ACCESS-CM2_ssp370_2021-2040.tif",
    file = tempfile()
  ),
  ssp585 = list(
    source = "https://geodata.ucdavis.edu/cmip6/5m/ACCESS-CM2/ssp585/wc2.1_5m_tmax_ACCESS-CM2_ssp585_2021-2040.tif",
    file = tempfile()
  )
)
```

```{r}
for (i in sources) {
  curl::curl_download(i$source, destfile = i$file, quiet = TRUE)
}

beepr::beep(1)
```

## Importing the data

```{r}
band <- "tmax06"
```

```{r}
for (i in names(sources)) {
  sources[[i]] <- c(
    sources[[i]],
    ras = 
      sources[[i]]$file |> 
      raster::stack() |> 
      magrittr::extract2(band)
  )
}
```

## Plotting the data

```{r}
#| eval: false

sa_shape <- 
  spData::world |>
  dplyr::filter(continent == "South America")
```

```{r}
#| eval: false

croped_ras <- 
  sources$ssp126$ras |> 
  raster::crop(raster::extent(sa_shape)) |>
  raster::mask(sa_shape)
```

```{r}
#| eval: false

plot_ras <- 
  croped_ras |>
  raster::rasterToPoints() |>
  dplyr::as_tibble()
```

```{r}
plot_ras |>
  ggplot2::ggplot(ggplot2::aes(x = x, y = y, fill = !!as.symbol(band))) + 
  ggplot2::geom_sf(
    data = sa_shape, 
    color = NA, 
    fill = "gray", 
    inherit.aes = FALSE
  ) +
  ggplot2::geom_tile() +
  # ggplot2::scale_fill_distiller(palette = "YlOrRd", direction = 1) +
  viridis::scale_fill_viridis(option = "B", direction = 1) +
  ggplot2::theme(
    axis.title.x = ggplot2::element_blank(),
    axis.title.y = ggplot2::element_blank()
  ) +
  ggplot2::labs(fill = "")
```

```{r}
color_range <- list(
  min = sources[[1]]$ras@data@min,
  max = sources[[1]]$ras@data@max
)

for (i in names(sources)) {
  color_range$min <- ifelse(
    color_range$min < sources[[i]]$ras@data@min, 
    color_range$min, 
    sources[[i]]$ras@data@min
  )
  
  color_range$max <- ifelse(
    color_range$max > sources[[i]]$ras@data@max, 
    color_range$max, 
    sources[[i]]$ras@data@max
  )
}

color_range <- c(
  floor(color_range$min), 
  ceiling(color_range$max)
)

color_range
```

To do: **Figure 10**. Surface air temperature differences (°C) for late 21st century 2081–2100. Minus 1995–2014 of the corresponding historical ensemble member for the non-interactive climate model. (a) SSP1-2.6; (b) SSP2-4.5; (c) SSP4-6.0; (d) SSP5-8.5. [@nazarenko2022].

```{r}
plot_ssp <- function(ssp, sources, color_range) {
  checkmate::assert_choice(ssp, c("ssp126", "ssp245", "ssp370", "ssp585"))
  checkmate::assert_list(sources)
  checkmate::assert_set_equal(names(sources[[1]]), c("source", "file", "ras"))
  checkmate::assert_integerish(color_range, len = 2)
  
  world_shape <-spData::world
  band <- names(sources[[1]]$ras)
  
  plot_ras <- 
    sources[[ssp]]$ras |>
    raster::rasterToPoints() |>
    dplyr::as_tibble()
  
  breaks <- c(color_range[1], 0, color_range[2])
  
  plot_ras |>
    ggplot2::ggplot(ggplot2::aes(x = x, y = y, fill = !!as.symbol(band))) + 
    ggplot2::geom_sf(
      data = world_shape, 
      color = NA, 
      fill = "gray", 
      inherit.aes = FALSE
    ) +
    ggplot2::geom_tile() +
    ggplot2::scale_fill_distiller(
      palette = "RdBu", 
      direction = -1,
      breaks = breaks,
      labels = as.character(breaks),
      limits = color_range
    ) +
    # ggplot2::scale_fill_gradient2(
    #   low = "purple4", 
    #   high = "darkred", 
    #   mid = "white", 
    #   midpoint = 0,
    #   limits = color_range
    # ) +
    # viridis::scale_fill_viridis(
    #   option = "B", 
    #   direction = 1,
    #   limits = color_range
    # ) +
    ggplot2::theme(
      axis.title.x = ggplot2::element_blank(),
      axis.title.y = ggplot2::element_blank()
    ) +
    ggplot2::labs(
      subtitle = toupper(ssp),
      fill = ""
    )
}
```

```{r}
color_range <- c(-60, 60)
```

```{r}
plot_ssp126 <- plot_ssp("ssp126", sources, color_range)

plot_ssp126
```

```{r}
plot_ssp245 <- plot_ssp("ssp245", sources, color_range)

plot_ssp245
```

```{r}
plot_ssp370 <- plot_ssp("ssp370", sources, color_range)

plot_ssp370
```

```{r}
plot_ssp585 <- plot_ssp("ssp585", sources, color_range)

plot_ssp585
```

```{r}
ggpubr::ggarrange(
  plot_ssp126, 
  plot_ssp245,
  plot_ssp370,
  plot_ssp585,
  ncol = 2, 
  nrow = 2,
  align = "v",
  legend = "bottom",
  common.legend = TRUE
)
```

## References {.smaller}

::: {style="font-size: 0.75em;"}
In accordance with the [American Psychological Association (APA) Style](https://apastyle.apa.org/), 7th edition.
:::

::: {#refs style="font-size: 0.75em;"}
:::
